---
title: "Day3_Practical1"
author: "Valentina Peona"
date: "2024-06-12"
output: html_document
---

# Day 3 - Practical 1

Model prior distributions for split time, population sizes, migration rate and ancestral population size.

## Simulate distributions for the parameters

T_split: we know that these populations have been separated by at least 1,000 generations and a
major environmental change between these two locations happened around 8,000 generations ago.

This prior can be modeled as a normal distribution with mean around 1000 and 8000 will be on the 97.5 percentile of the distribution.

N1/N2: we know from capture-recapture data that population 1 is approximately 30 times larger
than population 2; we also don't have a clear intuition of each value but from previous findings we expect values for population 1 to be between 50,000 and 200,000.

The priors for N1 and N2 are connected, therefore a normal distribution for N1 will just be scaled by 30 times for N2.

We expect either complete isolation after the split (rate=0) or pervasive migration (rate=0.1).

This prior can modeled as a bimodal distribution.

Nanc: we know this value with high confidence to be around 7,000,000

This prior can be modeled as a normal distribution with mean around 7,000,000 and a small standard deviation.


```{r}
# simulate distributions at least 500 times for each parameter
times = 500

migration = sample(x = c(0,1), size = times, prob = c(0.5,0.5), replace = TRUE)
Nanc = rnorm(times, mean = 7000000, sd = 10000)
N1 = rnorm(times, mean=125000, sd=25000)
N2 = N1/30

# sample the distribution of T_split until I get 500 values higher than 1000
Tsplit = rnorm(times, mean = 1000, sd = 3000)
Tsplit = Tsplit[Tsplit >= 1000]
diff = times - length(Tsplit)
while(diff > 0){
  
  new = rnorm(diff, mean = 1000, sd = 3000)
  new = new[new >= 1000]
  Tsplit = c(Tsplit, new)
  diff = times - length(Tsplit)
  
}
```

## Plot the distributions

```{r}
# Time of the split
hist(Tsplit, main = "Time of split", xlab = "Time", col = "lightblue", breaks = 20)

# Population sizy of N1, N2 and Nanc
hist(N1, main = "N1", xlab = "Population size", col = "lightblue", breaks = 20)
hist(N2, main = "N2", xlab = "Population size", col = "lightblue", breaks = 20)
hist(Nanc, main = "Nanc", xlab = "Population size", col = "lightblue", breaks = 20)

# Migration rate
hist(migration, main = "Migration rate", xlab = "Rate", col = "lightblue", breaks = 20)
```

## Run msprime

Assume a mutation rate of 3.5e-9, a recombination rate of 8.4e-9, and a sequence length of 1,000 bp.

  


