---
title: "Day3_Practical2"
author: "Valentina Peona"
date: "2024-06-12"
output: html_document
---

# Find the posterior probability distribution for parameters

All of this is done to understand if the demographic model used to simulate the data is good to explain the pattern we see in the data.

```{r}
# read the simulated data
sims <- read.csv("mosquito-task2.csv", head=T)
```

Plot the prior distributions

```{r}
# check prior distributions
#x11()
par(mfrow=c(2,2))
hist(sims$N1)
hist(sims$N2)
hist(sims$T_split)
hist(sims$MigRate)
```

Clean the data and find the summary statistics that correlate with the split time

```{r}
# remove simulations with NaN for some summary stats!
sims <- sims[complete.cases(sims),]

# find useful summary stats which correlate with T_split
cat("correlation of Fst with T_split\n")
cor(sims$Fst, sims$T_split)
cat("correlation of dxy with T_split\n")
cor(sims$dxy, sims$T_split)
cat("correlation of segsites1 with T_split\n")
cor(sims$segsites1, sims$T_split)
cat("correlation of segsites2 with T_split\n")
cor(sims$segsites2, sims$T_split)
cat("correlation of pi1 with T_split\n")
cor(sims$pi1, sims$T_split)
cat("correlation of pi2 with T_split\n")
cor(sims$pi2, sims$T_split)
cat("correlation of tajima1 with T_split\n")
cor(sims$tajima1, sims$T_split)
cat("correlation of tajima2 with T_split\n")
cor(sims$tajima2, sims$T_split)
```

Load the table with the observed summary statistics

```{r}
# load observed summary stats
obs <- read.csv("mosquito-observed.csv", head=T)
```

Check if the simulated retained summary stats contain the observed one

```{r}
# check if simulated retained summary stats contain the observed one
cat("Fst\n")
quantile(sims$Fst); cat("observed:", obs$Fst)
cat("\n\nsegsites1\n")
quantile(sims$segsites1); cat("observed:", obs$segsites1)
cat("\n\nsegsites2\n")
quantile(sims$segsites2); cat("observed:", obs$segsites2)
```

Scale the observed and simulated retained summary statistics

scaling must be done so that everything is comparable to the tolerance level

```{r}
# merge obs with retained sims to scale them: Fst, segsites1 and segsites2
sumstats <- scale(rbind(obs[c(1,3,4)],sims[,c(5,7,8)]))
```

Run the ABC analysis

```{r}
library(abc)

# the first row is given as target because it is the observed data
#est <- abc(target=sumstats[1,], param=sims$T_split, sumstat=sumstats[-1,], tol=0.05, method="rejection")
est = abc(target=sumstats[1,], param=sims[,c(1,2,3)], sumstat=sumstats[-1,], tol=0.05, method="rejection")
```

Plot the results

```{r}
# check distances in the acceptance region
hist(est$dist)
# whatever is falling before the vertical line are the values accepted
abline(v=max(est$dist[which(est$region)]), lty=2)

# posterior
#x11()
par(mfrow=c(2,2))
hist(est$unadj.values, freq=FALSE, xlim=range(sims$T_split), col=rgb(0,0,1,1/4), main="Posterior probability", xlab="Split time")
# MAP
map <- mean(est$unadj.values)
abline(v=map, lty=2)

# confidence intervals
hpd <- quantile(x=est$unadj.values, probs=c(0.025,0.975))
abline(v=hpd, lty=3)

# prior
hist(sims$T_split, freq=FALSE, xlim=range(sims$T_split), col=rgb(1,0,0,1/4))
```
